Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations

### 2021-03-26 papers/intro_paper/results_lvc.tex
changes in Sec. IV

The LIGO-Virgo Collaboration recently released their testing GR
catalogue containing results for events observed
during O3a~\cite{Abbott:2020jks}. For the test that we here present, the results shown in Ref.~\cite{Abbott:2020jks} only include events which pass a threshold for the median detector-frame total mass $\geq 90 \Mo$ and SNRs in the pre- and
post-merger regions $\geq 8$~\footnote{The pre- and post-merger regions of the signal are identified from the signal's power before and after the signal reaches the peak's amplitude, which is determined from the maximum of the likelihood function obtained with the parameter-estimation analysis. The SNR values are listed in Table IV of Ref.~\cite{Abbott:2020jks}}. 
%
The SNR threshold ensures that the signal contains sufficient
information in both the inspiral and merger stages to break the
degeneracy between the binary's total mass and the non-GR deviations
$(\df{220}, \dtau{220})$. Such strong degeneracy is present for
low-SNR events with negligible higher-modes and for which only the
post-merger is detectable rendering the measurement of $(\df{220},
\dtau{220})$ impossible for those cases \rb{(see Appendix~\ref{sec:correlation} for more details)}. On the other hand the
total mass threshold $\geq 90 \Mo$ was employed due to the fact that
this analysis is computationally expensive, and also because we
expected these events to be the most promising for ringdown
studies. However, since the SNR threshold alone should be sufficient
for the analysis, for this paper we run the test on all the events
listed in Ref.~\cite{Abbott:2020jks} that have SNRs in the pre- and
post-merger regions $\geq 8$, without imposing any mass threshold.

@@

However, \rb{in most non-GR theories,} the deviations parameters $(\df{220},\dtau{220})$ depend in general on the source's parameters, so \rb{if GR were to be wrong, one should expect their value to vary} across the GW signals observed by LIGO and Virgo.
As described in Ref.~\cite{Abbott:2020jks}, we can relax the assumption of a constant deviation across all events by using the hierarchical inference technique originally proposed in Refs.~\cite{Zimmerman:2019wzo,Isi:2019asy}. The general idea behind this technique is to assume that the non-GR parameters $(\df{220},\dtau{220})$ are drawn from a common underlying distribution, whose properties can be inferred from the population of events. Following~\cite{Zimmerman:2019wzo,Isi:2019asy,Abbott:2020jks} we model the population distribution with a Gaussian  $\mathcal{N}(\mu,\sigma)$ of unknown mean $\mu$ and standard deviation $\sigma$. Under those assumptions, the goal is then to directly measure a posterior distribution $P(\mu, \sigma |  \{d_j\}) $ for $\mu$ and $\sigma$ from a joint analysis of all the GW events. If GR is correct, then this posterior should be consistent with $\mu=0$ and $\sigma=0$. From Bayes' theorem it follows that~\cite{Isi:2019asy}:
%
\begin{equation}
P(\mu, \sigma |  \{d_j\}) \propto P(\mu,\sigma)\prod _{j=1}^N P(d_j |\, \mu, \sigma) \,,
\end{equation}
%
where $P(\mu,\sigma)$ is the prior (also known as hyperprior) on ($\mu,\sigma$), and $P(d_j | \mu, \sigma) $ can be written in terms of the individual likelihoods of a given non-GR parameter $\xi_{\text{nGR}}$ using~\cite{Isi:2019asy}
%
\begin{equation}
P(d_j | \mu, \sigma) = \int P(d_j |\, \xi_{\text{nGR}})  P(\xi_{\text{nGR}} |\, \mu, \sigma) \,d\xi_{\text{nGR}} \,.
\end{equation}
%
Here $P(\xi_{\text{nGR}} | \,\mu, \sigma) = \mathcal{N}(\mu,\sigma)$ by construction and $P(d_j |\, \xi_{\text{nGR}})$ is the likelihood for the parameter $\xi_{\text{nGR}}$ for a given event $d_j$ that is computed from the standard parameter estimation analysis. After obtaining $P(\mu, \sigma |  \{d_j\})$, we can then infer a population distribution for a given non-GR parameter $\xi_{\text{nGR}}$ using~\cite{Isi:2019asy}:
%
\begin{equation}\label{hier_combine}
P(\xi_{\text{nGR}} | \{d_j\}) = \int P(\xi_{\text{nGR}} |\, \mu, \sigma)\,P(\mu, \sigma |  \{d_j\})\,d\mu\,d\sigma \,.
\end{equation}
%
Notice that, if we fix $\sigma=0$, this approach is equivalent to assume that all events share the same non-GR parameter $\xi_{\text{nGR}} =\mu$ and Eq.~\eqref{hier_combine} reduces to the joint likelihood~\cite{Zimmerman:2019wzo}. In practice, we use the \texttt{stan}-based code~\cite{stan} developed and used in Refs.~\cite{Isi:2019asy,Abbott:2020jks} to obtain $P(\mu, \sigma |  \{d_j\}) $ and compute $P(\xi_{\text{nGR}} | \{d_j\})$.

>>>>

The LIGO-Virgo Collaboration recently released their testing GR
catalogue containing results for events observed
during O3a~\cite{Abbott:2020jks}. For the test that we here present, the results shown in Ref.~\cite{Abbott:2020jks} only include events which pass a threshold for the median detector-frame total mass $\geq 90 \Mo$ and SNRs in the pre- and
post-merger regions $\geq 8$~\footnote{The pre- and post-merger regions of the signal are identified from the signal's power before and after the signal reaches the peak's amplitude, which is determined from the maximum of the likelihood function obtained with the parameter-estimation analysis. The SNR values are listed in Table IV of Ref.~\cite{Abbott:2020jks}}. 
%
The SNR threshold ensures that the signal contains sufficient
information in both the inspiral and merger stages to break the
degeneracy between the binary's total mass and the non-GR deviations
$(\df{220}, \dtau{220})$. Such strong degeneracy is present for
low-SNR events with negligible higher-modes and for which only the
post-merger is detectable rendering the measurement of $(\df{220},
\dtau{220})$ impossible for those cases (see Appendix~\ref{sec:correlation} for more details). On the other hand the
total mass threshold $\geq 90 \Mo$ was employed due to the fact that
this analysis is computationally expensive, and also because we
expected these events to be the most promising for ringdown
studies. However, since the SNR threshold alone should be sufficient
for the analysis, for this paper we run the test on all the events
listed in Ref.~\cite{Abbott:2020jks} that have SNRs in the pre- and
post-merger regions $\geq 8$, without imposing any mass threshold.

@@

However, in most non-GR theories, the deviations parameters $(\df{220},\dtau{220})$ depend in general on the source's parameters, so if GR were to be wrong, one should expect their value to vary across the GW signals observed by LIGO and Virgo.
As described in Ref.~\cite{Abbott:2020jks}, we can relax the assumption of a constant deviation across all events by using the hierarchical inference technique originally proposed in Refs.~\cite{Zimmerman:2019wzo,Isi:2019asy}. The general idea behind this technique is to assume that the non-GR parameters $(\df{220},\dtau{220})$ are drawn from a common underlying distribution, whose properties can be inferred from the population of events. Following~\cite{Zimmerman:2019wzo,Isi:2019asy,Abbott:2020jks} we model the population distribution with a Gaussian  $\mathcal{N}(\mu,\sigma)$ of unknown mean $\mu$ and standard deviation $\sigma$. Under those assumptions, the goal is then to directly measure a posterior distribution $P(\mu, \sigma |  \{d_j\}) $ for $\mu$ and $\sigma$ from a joint analysis of all the GW events. If GR is correct, then this posterior should be consistent with $\mu=0$ and $\sigma=0$. From Bayes' theorem it follows that~\cite{Isi:2019asy}:
%
\begin{equation}
P(\mu, \sigma |  \{d_j\}) \propto P(\mu,\sigma)\prod _{j=1}^N P(d_j |\, \mu, \sigma) \,,
\end{equation}
%
where $P(\mu,\sigma)$ is the prior (also known as hyperprior) on ($\mu,\sigma$), and $P(d_j | \mu, \sigma) $ can be written in terms of the individual likelihoods of a given non-GR parameter $\xi_{\text{nGR}}$ using~\cite{Isi:2019asy}
%
\begin{equation}
P(d_j | \mu, \sigma) = \int P(d_j |\, \xi_{\text{nGR}})  P(\xi_{\text{nGR}} |\, \mu, \sigma) \,d\xi_{\text{nGR}} \,.
\end{equation}
%
Here $P(\xi_{\text{nGR}} | \,\mu, \sigma) = \mathcal{N}(\mu,\sigma)$ by construction and $P(d_j |\, \xi_{\text{nGR}})$ is the likelihood for the parameter $\xi_{\text{nGR}}$ for a given event $d_j$ that is computed from the standard parameter estimation analysis. After obtaining $P(\mu, \sigma |  \{d_j\})$, we can then infer a population distribution for a given non-GR parameter $\xi_{\text{nGR}}$ using~\cite{Isi:2019asy}:
%
\begin{equation}\label{hier_combine}
P(\xi_{\text{nGR}} | \{d_j\}) = \int P(\xi_{\text{nGR}} |\, \mu, \sigma)\,P(\mu, \sigma |  \{d_j\})\,d\mu\,d\sigma \,.
\end{equation}
%
Notice that, if we fix $\sigma=0$, this approach is equivalent to assume that all events share the same non-GR parameter $\xi_{\text{nGR}} =\mu$ and Eq.~\eqref{hier_combine} reduces to the joint likelihood~\cite{Zimmerman:2019wzo}. In practice, we use the \texttt{stan}-based code~\cite{stan} developed and used in Refs.~\cite{Isi:2019asy,Abbott:2020jks} to obtain $P(\mu, \sigma |  \{d_j\}) $ and compute $P(\xi_{\text{nGR}} | \{d_j\})$.
